\begin{abstract}

%Then, we start in the abstract with the story that a network deployment and corresponding cost structure 
%in a dense urban area should look very different from that in a rural area.  Mention that white 
%spaces allow the degree to which such scalability can truly be achieved due to the higher degree of 
%propagation, allowing greater levels of aggregation.  This is in direct contrast to urban areas which 
%try to maximize spatial reuse.   Then, moving to the introduction, the focus is not on metropolitan 
%deployment but rather rural deployment.  You can cite some announcements by the FCC or federal 
%government of the intention of the white space bands (rural areas).  You can then say that we perform 
%a measurement study which considers these propagation characteristics and observed spectrum utility 
%to consider the amount of cost savings a rural network could achieve with the use of white spaces. 
%Then, all of the huge gains you show now really support this story. 
%
Wireless Mesh Network is an efficient and low cost solution for large-scale Internet connectivity.
A network deployment and corresponding cost structure in a dense urban area should look very different
from that in a rural area. FCC intends to improve the coverage of rural area in U.S. by reapportioning 
spectrum for data communication with far greater range in lower carrier frequencies. These white 
spaces allow the degree to which such scalability can truly be achieved due to the higher degree of 
propagation, allowing greater levels of aggregation. The large propagation range is direct contrast to 
urban areas which try to maximize spatial reuse. Thus, leveraging the spectrum across diverse
population densities and finding the proper bands become an urgent task for the deployment of large-scale 
mesh access networks. In this work, we measure the spectrum utility in DFW area and propose a measurement 
driven band selection framework Multiband Access Point Estimation (MAPE) which explores the white space
bands with in-field spectrum utility measurements telling the number of access points required of an area 
with channels in multiple bands. In doing so, we find the gains of white space bands application over 
existing WiFi only application in reducing the number of access points up to 1650\% in sparse 
rural area, 660\% in rural area, and 412\% in sparse urban area. Moreover, due to the heavy used white 
space bands, the spacial reusable WiFi bands gains 6.25\% reducing the number of access points in 
urban area, and 6.6\% in downtown area. We use our MAPE framework numerical analyze band combination 
in typical urban area and the result shows the more channels in white space band, the less access points 
are required for an arbitrary areas with the same number of channels in total.  


%To deploy new wireless mesh network, knowing the existing signals in the air is a 
% must. The existing signal information is important for all the problems in wireless mesh network, such as
% gateway deployment, channel assignment, and routing. Many work has been done to approach the interference model.
% However, there is no model fit for all metropolitan cities. Moreover, FCC regulations have reapportioned spectrum for 
% data communication with far greater range than WiFi in lower carrier frequencies. 
% Thus, leveraging the spectrum utility and finding the proper bands become an
%  urgent task for metropolitan cities who want to deploy wireless mesh network.
% In this work, we measure the spectrum activities in DFW area and propose a measurement 
% driven band selection framework Multiband Access Point Estimation (MAPE) which explore the white space
% bands with in-field spectrum utility telling the number of access points required to serve an area
% with channels in different bands.
% In doing so, we find gains of white space bands application over existing multi-channel, 
% multi-radio application in reducing the number of access points up to 1650\% in sparse 
% rural area, 660\% in rural area, and 412\% in sparse urban area.
% Moreover, due to the heavy used white space band, the spacial reusable WiFi bands gains 6.25\% reducing 
% the number of access points in urban area, and 6.6\% in downtown area.
% Moreover, we use our MAPE framework numerical analyze band combination in typical urban area and the result
% shows with the same number of channels, the more channels in white space band, the less access points needs for
% a candidate areas.  
\end{abstract}
