\section{Related Work}
\label{sec:related}

% Inter-network vs Intra-network interference
Many efforts have been put in wireless network deployment in WiFi bands~\cite{he2008optimizing,ramachandran2006interference,akyildiz2006next}.
These works focus on solving gateway placement, channels assignment, routing 
problems in wireless network issues to reduce the interference generated inside the network.
However, wireless interference is not only from a new deployment network but also from the
existing wireless device. Unfortunately, few of the works in network deployment notice the 
inter-network interference. There are some cognitive radio works discuss the inter-network
interference, but most of them focus on point to point communication other than network~\cite{cabric2004implementation}.

% White Space
As FCC adopt white space bands for wireless communication, questions come out 
for wireless network deployment~\cite{fccwhitespace}. The white spaces bring two
concerns: large propagation range and existing Inter-network interference from 
TV station and other device~\cite{cui2013leveraging,bahl2009white}.
Previous works fails to discover the white spaces benefit and drawback in network deployment~\cite{akyildiz2005wireless}.

% FIXME
Brand new white spaces problems are new research hot spot. Follow FCC's policy, white space capacity 
of  continental USA is quantified in~\cite{harrison2010much}. White space is targeting on opportunistic 
medium access for data transmission and the opportunity is experimental verified indoor~\cite{obregon2010experimental}.
However, the best application of white space is to serve rural area with its long propagation range
other than indoor application. Our target is to quantify the opportunity of outdoor backhual network.

 % Network propagation difference
There are some works talked about the propagation variation in both WiFi network and white space network.
In~\cite{robinson2010deploying}, Robinson model the propagation variation in the same band for terrain 
domain. Rohan bring a databased-driven SenseLess framework for operating a white space network with 
database of primary user location, channel occupation~\cite{murty2012senseless}. 
However, as opportunistic medium access, the most important application of white space bands are 
focus on or related to the time domain. It is impossible to build a mathematical model of white 
space bands due to the variation of FCC rules 
in multiple metropolitan. Measurement is the best way to tell the current state of bands utility.
We propose the measurement framework with in-field data for wireless network deployment band selection.


 


